---
title: "Data cleaning"
author: "Jae Yeon Kim"
output:
html_document: 
  toc: true
  theme: united
---

## 0. Setup 

I tweaked the global option of the R Markdown to enlarge figures produced by ggplot2.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, 
                      echo = FALSE, warning = FALSE, message = FALSE) # global setting for enlarging image size
```

```{r}

# Clean up the environment

rm(list = ls())

# Import libraries (adapted from this link: https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
        tidyverse, # for the tidyverse framework
        naniar, # for missing data
        mice, # for imputation
        ggpubr, # for arranging ggplots   
        ggthemes, # for fancy ggplot themes
        BaylorEdPsych # testing patterns among missing values
)

```

### 1. Importing data 

Survey participants: 4,435 
Questions or features: 131 

```{r}

df <- read_csv("/home/jae/validating-two-linked-fates/raw_data/sub-data.txt")

head(df)
```

### 2. Subsetting 

```{r}
df <- df %>%
  dplyr::select(ETHNICITY, # ETHNICITY weight: 1 - White, 2 - Latino, 3 - Black, 4 - Asian, 5 - Other/Refuesed/No answer
                Q16d1, # Linked fate:  Do you agree or disagree - What happens to (Latinos/ whites/ blacks and African Americans / Asians/ Pacific Islanders/ American Indians) in this country will have something to do with what happens in my life (1 - Agree strongly, 4 - Disagree strongly)
                Q16d2, # Linked progress: Do you agree or disagree - When things get better for (Latinos/ whites/ blacks and African Americans/ Asians/ Pacific Islanders/ American Indians) in this country, then things will get better for me
                Q16d3, # Linked hurt: Do you agree or disagree - When things get worse for (Latinos/ whites/ blacks and African Americans/ Asians/ Pacific Islanders/ American Indians) in this country, then things will get worse for me

                Q32a, # Generally speaking in politics do you think of yourself as a . . . Partisanship: 1 - GOP, 2 - Democrat, 3 - Neither 
                WEIGHT, # Survey weight
                AGE,
                Gender, 
                party, # part of registration 
                Q40, # Foreign born status: 1 - Born in U.S., 2 - Born in another country 
                Q16a) # Education level: 1 - lowest, 8 - highest 

```

### 3. Inspeacting missing values 

- Inspecing missing data. 
- Percentage of missing observations: only 1.8%

```{r}
vis_miss(df)
```

### 4. Rescaling and renaming variables 

```{r}

# Create a function to rescale linked fate related questions 
rescale <- function(data){
  data %>% 
    as.character() %>%
      recode("4" = "-2", 
        "3" = "-1", 
        "2" = "1", 
        "1" = "2") %>%
    as.factor()
}

# Rescaling and renaming variables 
df <- df %>%
  mutate(Q16d1 = rescale(Q16d1),
         Q16d2 = rescale(Q16d2), 
         Q16d3 = rescale(Q16d3)) %>%
  rename("linked_fate"= "Q16d1",
         "linked_progress" = "Q16d2",
         "linked_hurt" = "Q16d3",
         "for_born" = "Q40",
         "edu_level" = "Q16a",
         "party_id" = "Q32a",
         "party_reg" = "party",
         "race" = "ETHNICITY")

```
### 5. Recoding values

I dropped 21 partipants (less than 1% of the total survey participants) because we can't use Q16c variable to identify their race. 

```{r}
table(df$race) # We only include those responded 1-5 for Q16c variable 

df <- subset(df, race >= 1 & race <= 5)

length(df$race)

df$race <- df$race %>%
  as.character() %>%
  recode("1" = "White", 
        "2" = "Latinx", 
        "3" = "Black", 
        "4" = "Asian", 
        "5" = "Other") %>%
  as.factor()

```

```{r}
write.csv(df, "/home/jae/validating-two-linked-fates/raw_data/processed_survey.csv")
```

